### 2022-08-30 preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d74d2d8fe84f81e9431ea1c135d8cb274a599339 🚀
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@e95f073e8afe0ab4deb3916811d8ca7d3cf74bd9 🚀

\medskip
\begin{flushleft}
{\small
%
Norma\,Forero Rocio Munoz%
%
\,\textsuperscript{1,2,‡}, %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}, %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}, %
Gregory F.\,Albery%
%
\,\textsuperscript{5}, %
Daniel J.\,Becker%
%
\,\textsuperscript{6}, %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}
\vskip 1em
\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
Epidemiology and Public Health Laboratory, School of Veterinary Science,
Massey University, New Zealand; \textsuperscript{4}\,Paul G. Allen
School for Global Health, Washington State University, Pullman, WA,
United States; \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA; \textsuperscript{6}\,Department of
Biology, University of Oklahoma, Norman, OK,
USA; \textsuperscript{7}\,Department of Biology, Georgetown University,
Washington, DC,; \textsuperscript{8}\,Center for Global Health Science
and Security, Georgetown University Medical Center, Washington, DC,
USA; \textsuperscript{9}\,Department of Microbiology and Immunology,
Georgetown University Medical Center, Washington, DC, USA\\
\textsuperscript{‡}\,These authors contributed equally to the work\\
\vskip 1em
\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
}
\end{flushleft}

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{66} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{67} v1.6.12
with ModelFinder\textsuperscript{68} ultrafast bootstrap
approximation\textsuperscript{69} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{70} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.

>>>>

\medskip
\begin{flushleft}
{\small
%
Norma R\,Forero-Muñoz%
%
\,\textsuperscript{1,2,‡}, %
Renata L.\,Muylaert%
%
\,\textsuperscript{3}, %
Stephanie N.\,Seifert%
%
\,\textsuperscript{4}, %
Gregory F.\,Albery%
%
\,\textsuperscript{5}, %
Daniel J.\,Becker%
%
\,\textsuperscript{6}, %
Colin J.\,Carlson%
%
\,\textsuperscript{7,8,9,‡}, %
\href{https://orcid.org/0000-0002-0735-5184}{Timothée\,Poisot}%
%
\,\textsuperscript{1,2,‡}
\vskip 1em
\textsuperscript{1}\,Université de Montréal; \textsuperscript{2}\,Québec
Centre for Biodiversity Sciences; \textsuperscript{3}\,Molecular
Epidemiology and Public Health Laboratory, School of Veterinary Science,
Massey University, New Zealand; \textsuperscript{4}\,Paul G. Allen
School for Global Health, Washington State University, Pullman, WA,
United States; \textsuperscript{5}\,Department of Biology, Georgetown
University, Washington, DC, USA; \textsuperscript{6}\,Department of
Biology, University of Oklahoma, Norman, OK,
USA; \textsuperscript{7}\,Department of Biology, Georgetown University,
Washington, DC,; \textsuperscript{8}\,Center for Global Health Science
and Security, Georgetown University Medical Center, Washington, DC,
USA; \textsuperscript{9}\,Department of Microbiology and Immunology,
Georgetown University Medical Center, Washington, DC, USA\\
\textsuperscript{‡}\,These authors contributed equally to the work\\
\vskip 1em
\textbf{Correspondance to:}\\
Timothée Poisot --- \texttt{timothee.poisot@umontreal.ca}\\
}
\end{flushleft}

@@

To next represent phylogeography of betacoronaviruses in bats, we
aggregated and analyzed betacoronavirus sequence data. We used the
following query to pull all \emph{Betacoronavirus} sequence data from
the GenBank Nucleotide database except SARS-CoV-2;
(``Betacoronavirus''{[}Organism{]} OR betacoronavirus{[}All Fields{]})
NOT (``Severe acute respiratory syndrome coronavirus 2''{[}Organism{]}
OR sars-cov-2{[}All Fields{]}). We added a single representative
sequence for SARS-CoV-2 and manually curated to remove sequences without
the RNA-dependent RNA polymerase (RdRp) sequence or that contained words
indicating recombinant or laboratory strains including ``patent'',
``mutant'', ``GFP'', and ``recombinant''. We filtered over-represented
taxa including betacoronavirus 1, hCoV-OC43, Middle East respiratory
syndrome coronavirus, Murine hepatitis virus, and hCoV-HKU1. Curated
betacoronavirus RdRp sequences were then aligned using
MAFFT\textsuperscript{67} v1.4.0 (Algorithm FFT-NS-2, Scoring matrix
200PAM / k=2, gap open penalty 1.53m offset value 0.123) and a maximum
likelihood tree reconstructed in IQ-TREE\textsuperscript{68} v1.6.12
with ModelFinder\textsuperscript{69} ultrafast bootstrap
approximation\textsuperscript{70} with a general time reversible model
with empirical base frequencies and the 5-discrete-rate-category
FreeRaye model of nucleotide substitution (GTR+F+R5).

We first tested the hypothesis that hotspots of viral diversification
would track hotspots of bat diversification. To do so, we plotted the
number of known bat hosts (specifically only those included in the
phylogeny, so there was a 1:1 correspondence between data sources)
against the ``mean evolutionary distinctiveness'' of the associated
viruses. To calculate this, we derived the fair proportions evolutionary
distinctiveness\textsuperscript{71} for each of the viruses in the tree,
then averaged these at the bat species level, projected these values
onto their geographic distributions, and averaged across every bat found
in a given pixel. As such, this can be thought of as a map of the mean
evolutionary distinctiveness of the known viral community believed to be
associated with a particular subset of bats present.
